---
title: "Untitled"
author: "Jacob Gerzenshtein"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cancer Cell Work

Importing data and rearranging the B and M conditions to Boolean 

```{r}
wisc.df <- read.csv("WisconsinCancer.csv", head = TRUE)

View(wisc.df)

wisc.data <- as.matrix(wisc.df[ ,3:32])

row.names(wisc.data) <- wisc.df$id

diagnosis <- as.numeric(wisc.df$diagnosis == "M")

table(diagnosis)

table(wisc.df$diagnosis)
```

# Q1: How many observations are in this data set? 

```{r}
row <- nrow(wisc.df)
column <- ncol(wisc.df)

row 

column

total <- row * column

total 

```
There are 569 observations in this data set (569 rows, 33 columns)


# Q2: How many variables/features in the data are suffixed with _mean?

```{r}
# The column names are not counted as the data, have to specify that you want to 
# search them
z <- colnames(wisc.data)
grep("_mean", z, value = TRUE)
```

There are 10 variables in the data suffixed with _mean


# Q3: How many of the observations have a malignant diagnosis? 

```{r}
sum(diagnosis)
```

There are 212 Malignant cases 

# Checking the mean and standard deviation of the features 

```{r}
colMeans(wisc.data)

apply(wisc.data, 2, sd)
```

# Using PCA now 

```{r}
x <- scale(wisc.data, scale = TRUE)

wisc.pr <- prcomp(x)

summary(wisc.pr)
```


# Q4: From results, what proportion of the original variance is captured by the first principal components (PC1) ?

44.27%

# Q5: How many principal components (PCs) are required to describe at least 70% of the original variance in the data? 

```{r}
sum(44.27 + 18.97 + 9.4)
```

3 PC's needed to get at least 70% 


# Q6: How many PC's needed to describe at least 90% of the original variance in the data?

7 PC's needed to get at least 90%

# Creating a biplot 

```{r}
biplot(wisc.pr)
```

This plot is difficult to understand - everything is cluttered and clustered, with the observations being clustered together in black in one corner and the variables being clustered together in red in the other corner 

# A different plot

```{r}
plot(wisc.pr$x[ , c(1,2)], col  = c(diagnosis + 1), xlab = "PC1", ylab = "PC2")
```


```{r}
plot(wisc.pr$x[ , c(1,3)], col  = c(diagnosis + 1), xlab = "PC1", ylab = "PC2")
```



# Scree plot

```{r}
pr.var <- wisc.pr$sdev^2

pve <- pr.var / sum(pr.var)

par(mfrow = c(1,3))

barplot(pve, ylab = "Percent of Variance Explained", names.arg = paste0("PC", 1:length(pve)), las = 2, axes = FALSE) 

axis(2, at=pve, labels = round(pve,2)*100)

plot(pve, xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = "o")

plot(cumsum(pve), xlab = "Principal Component", yab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = "o")
```

## More PCA Work

```{r}
#scale the data
data.scaled <- scale(wisc.data)

#create the distance necessary for the hclust command
data.dist <- dist(data.scaled)

#create the hclust in orderto help creat the dendrogram 
wisc.hclust <- hclust(data.dist)

#create 4 clusters from the data 
wisc.hclust.clusters <- cutree(wisc.hclust, k = 9)

#plot the desired dendrogram 
plot (wisc.hclust)

abline(h = 20, col = "red")
```

Creating a table 

```{r}
table(wisc.hclust.clusters, diagnosis)
```

# Q 12: Better clusters? 

It would be better to have 9 clusters 

# Calling K means

```{r}
wisc.km <- kmeans(wisc.data, centers = 2, nstart = 20)

wisc.km
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method = "complete")

wisc.pr.hclust
```

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k = 4)

plot(wisc.pr.hclust)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)
```

