---
title: "BIMM 143 Networks"
author: "Jacob Gerzenshtein"
date: "May 29, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Network Analysis

Examining metagenomics of oceanic data using Cytoscape and R 

Loading the necessary packages

```{r}
install.packages( c("igraph", "RColorBrewer"))

source("https://bioconductor.org/biocLite.R")

biocLite("RCy3")
```

```{r}
library(RCy3)
library(igraph)
library(RColorBrewer)
```

Verifying the connection to Cytoscape 

```{r}
cwd <- demoSimpleGraph()
```

Configuring

```{r}
layoutNetwork(cwd, 'force-directed')

#Setting a layout 

possible.layout.names <- getLayoutNames(cwd)

layoutNetwork (cwd, possible.layout.names[1])

setVisualStyle(cwd, "Ripple")

#Testing the connection

ping(cwd)
```


Saving image

```{r}
saveImage(cwd,
          file.name = "demo",
          image.type = "png",
          h = 350)
```


The connection is established, now need to import and read the Data

```{r}
prok_vir_cor <- read.delim("virus_prok_cor_abundant.tsv", stringsAsFactors = FALSE)

head(prok_vir_cor)
```

Now using igraph to work with network and Cytoscape 

Converting the co-occurence of dataframe into network that can be sent to Cytoscape 

Creating an undirected network 

```{r}
g <- graph.data.frame(prok_vir_cor, directed = FALSE)

class (g)
```

Taking an initial look

```{r}
plot (g)
```

Inspecting

```{r}
V(g)

E(g)
```


Tuning the plot a little

```{r}
plot(g, vertex.size = 3, vertex.label = NA)
```

Reading the bacterial data 

```{r}
phage_id_affiliation <- read.delim("phage_ids_with_affiliation.tsv")

head(phage_id_affiliation)

bac_id_affi <- read.delim("prok_tax_from_silva.tsv")

head(bac_id_affi)
```


Creating the network that Cytoscape can use 



```{r}
genenet.nodes <- as.data.frame(vertex.attributes(g))

#Creating empty columns because not all have classifications

genenet.nodes$phage_aff <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$Tax_order <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$Tax_subfamily <- rep("not_class", nrow(genenet.nodes))

for (row in seq_along(1:nrow(genenet.nodes))) {
  if (genenet.nodes$name[row] %in% phage_id_affiliation$first_sheet.Phage_id_network) {
    id_name <- as.character(genenet.nodes$name[row])
    aff_to_add <- unique(subset(phage_id_affiliation,
                         first_sheet.Phage_id_network == id_name,
                         select = c(phage_affiliation,
                                    Tax_order,
                                    Tax_subfamily)))
    genenet.nodes$phage_aff[row] <- as.character(aff_to_add$phage_affiliation)
    genenet.nodes$Tax_order[row] <- as.character(aff_to_add$Tax_order)
    genenet.nodes$Tax_subfamily[row] <- as.character(aff_to_add$Tax_subfamily)

  }
}
```

For Prokaryotes
```{r}
genenet.nodes$prok_king <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$prok_tax_phylum <- rep("not_class", nrow(genenet.nodes))
genenet.nodes$prok_tax_class <- rep("not_class", nrow(genenet.nodes))

for (row in seq_along(1:nrow(genenet.nodes))) {
  if (genenet.nodes$name[row] %in% bac_id_affi$Accession_ID) {
    id_name <- as.character(genenet.nodes$name[row])
    aff_to_add <- unique(subset(bac_id_affi,
                         Accession_ID == as.character(genenet.nodes$name[row]),
                         select = c(Kingdom,
                                    Phylum,
                                    Class)))
    genenet.nodes$prok_king[row] <- as.character(aff_to_add$Kingdom)
    genenet.nodes$prok_tax_phylum[row] <- as.character(aff_to_add$Phylum)
    genenet.nodes$prok_tax_class[row] <- as.character(aff_to_add$Class)

  }
}
```

Adding the relevant network data for connections between organisms, then sending to Cytoscape via cyPlot

```{r}
genenet.edges <- data.frame(igraph::as_edgelist(g))
names(genenet.edges) <- c("name.1", "name.2")

genenet.edges$Weight <- igraph::edge_attr(g)[[1]]

genenet.edges$name.1 <- as.character(genenet.edges$name.1)
genenet.edges$name.2 <- as.character(genenet.edges$name.2)
genenet.nodes$name <- as.character(genenet.nodes$name)

ug <- cyPlot(genenet.nodes, genenet.edges)

```

Sending the network from R to Cytoscape 

Deleting current windows and establishing the connection

```{r}
cy <- CytoscapeConnection()

#deleting the current windows

deleteAllWindows(cy)
```

```{r}
cw <- CytoscapeWindow("Tara oceans",
                      graph = ug,
                      overwriteWindow = TRUE)
```

```{r}
displayGraph(cw)
layoutNetwork(cw)
fitContent(cw)
```

Formatting the plot

```{r}
families_to_colour <- unique(genenet.nodes$prok_tax_phylum)
families_to_colour <- families_to_colour[!families_to_colour %in% "not_class"]
node.colour <- RColorBrewer::brewer.pal(length(families_to_colour), "Set3")
```

Coloring Nodes

```{r}
setNodeColorRule(cw,
                 "prok_tax_phylum",
                 families_to_colour,
                 node.colour,
                 "lookup",
                 default.color = "#ffffff")
```

```{r}
displayGraph(cw)
layoutNetwork(cw)
fitContent(cw)
```

Changing shape of node to reflect viral vs prokaryotic 

```{r}
shapes_for_nodes <- c("DIAMOND")

phage_names <- grep("ph_",
                    genenet.nodes$name,
                    value = TRUE)

setNodeShapeRule(cw,
                 "label",
                 phage_names,
                 shapes_for_nodes)
```

```{r}
displayGraph(cw)

fitContent(cw)
```

Colouring by family 

```{r}
setDefaultNodeBorderWidth(cw, 5)

families_to_colour <- c("Podoviridae",
                        "Siphoviridae",
                        "Myoviridae")

node.colour <- RColorBrewer::brewer.pal(length(families_to_colour), "Dark2")

setNodeBorderColorRule(cw,
                       "Tax_subfamily",
                       families_to_colour,
                       node.colour,
                       "lookup",
                       default.color = "#000000")
displayGraph(cw)

fitContent(cw)
```

Setting up a layout to minimize node overlap

```{r}
getLayoutNames(cw)

getLayoutPropertyNames(cw, layout.name = "force-directed")

getLayoutPropertyValue(cw, "force-directed", "defaultSpringLength")

getLayoutPropertyValue(cw, "force-directed", "numIterations")

setLayoutProperties(cw,
                    layout.name = "force-directed",
                    list(defaultSpringLength = 20,
                        "numIterations" = 200))

layoutNetwork(cw,
              layout.name = "force-directed")

fitContent(cw)
```

