---
title: "Bioinformatics Class 7"
author: "Jacob Gerzenshtein"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Functions Again

Here we are going to revisit our function from class 6

```{r}
source("http://tinyurl.com/rescale-R")
```

Let's see if we can use this function


```{r}
rescale(1:10)
```


Looks good - let's break it 

```{r}
#rescale(c(1:10, "string"))
```

Let's try the new **rescale2** function

```{r}
#rescale2(c(1:10, "string"))
```

## Writing a NA checking function

Here we write a new function to check for NA's in two inputs

```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
which(is.na(x))
which(is.na(y))
```

```{r}
sum (is.na(x))

```

```{r}
is.na(x)
is.na(y)
```


```{r}
is.na(x) & is.na(y)
```



Let's out this together with x and y 

```{r}
sum(is.na(x) & is.na(y))
```

## The function:
```{r}
both_na <- function(x, y) {



sum(is.na(x) & is.na(y))

}
```

Test it

```{r}
both_na(x,y)
```


Breaking it: 

```{r}
x <- c(NA, NA, NA)
y1 <- c(1, NA, NA)
y2 <- c(1, NA, NA, NA)

both_na(x,y)
```

New function to create warning:

```{r}
both_na2 <- function (x,y) {
  
  
  if (length(x) != length(y)) {
    
    stop("Input x and y should be the same length")
    
  }
  
  sum(is.na(x) & is.na(y))
  
}
```

Test it again in console 


A new function to find number of matches and where they are:

```{r}
both_na3 <- function(x,y) {
  
  if (length(x) != length(y)) { 
    
    stop("Input x and y should be vectors of the same length")
    
    }
  
  na.in.both <- (is.na(x) & is.na(y))
  na.number <- sum(na.in.both)
  na.which <- which(na.in.both)
  
  message("Found", na.number, "NA's at position(s):", paste(na.which, collapse = ", "))
  
  
  return(list(number=na.number, which=na.which))
  
  
}
```

Try it: 

```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
both_na3(x,y)
```





## Follow along: a different function for gene intersection 


Sourcing the code 
```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
df1
df2

x <- df1$IDs
y <- df2$IDs


```

```{r}
x
y
```


Seeing where there are similarities 
```{r}
intersect (x,y)
x%in%y
y%in%x
```

Now access the genes with desired indices 

```{r}
x[x%in%y]
y[y%in%x]
```


Now put them together with **cbind()**

```{r}
cbind(x[x%in%y], y[y%in%x])
```


```{r}
gene_intersect <- function (df1, df2) {

cbind(df1[df1$IDs %in% df2$IDs, ],
      df2[df2$IDs %in% df1$IDs, "exp"])

}
```

```{r}
gene_intersect2(df1, df2)
```



```{r}
merge(df1, df2, by = "IDs")
```

