---
title: "BIMM 143 Differential Expression"
author: "Jacob Gerzenshtein"
date: "May 22, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Differential Expression Analysis 

Loading the necessary packages

```{r}
library(DESeq2)
```

Loading data files


```{r}
metaFile <- "GSE37704_metadata.csv"

countFile <- "GSE37704_featurecounts.csv"
```

Reading some of the data

```{r}
colData <- read.csv(metaFile, row.names = 1)
head(colData)
```

Importing count data

```{r}
countData <- read.csv(countFile, row.names = 1)
head(countData)
```

Need all the columns to match up so need to remove the first column of countData

```{r}
countData <- as.matrix(countData[ ,-1])
head(countData)
```

Getting rid of the entries that are just zeroes 

```{r}
countData <- countData[rowSums(countData) > 1, ]
head(countData)
```


Setting up DESeq2 pipeline and analysis 

```{r, cache=TRUE}
dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~condition)

dds <- DESeq(dds)
```


Seeing what the output is

```{r}
dds
```

Running analysis of the HoxA1 knockdown versus control siRNA

```{r}
res = results(dds, contrast = c("condition", "hoxa1_kd", "control_sirna"))

# reordering by p-value

res = res[order(res$pvalue), ]

summary(res)
```
 
Introducing ENTREZ as a source of data to annotate the existing output

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)

columns(org.Hs.eg.db)
```


Using the mapID function

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys = row.names(res),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")

res$entrez <- mapIds(org.Hs.eg.db,
                     keys = row.names(res),
                     column = "ENTREZID",
                     keytype = "ENSEMBL",
                     multiVals = "first")

res$name = mapIds(org.Hs.eg.db,
                  keys = row.names(res),
                  column = "GENENAME",
                  keytype = "ENSEMBL",
                  multiVals = "first")

head(res, 10)
```


## Using KEGG tools, and the gageData package

Installing necessary packages

```{r}
#source("http://bioconductor.org/biocLite.R")

#biocLite( c("pathview", "gage", "gageData"))
```

Loading necessary and relevant packages

```{r}
library(pathview)

library(gage)

library(gageData)
```

Reading Data

```{r}
data("kegg.sets.hs")
data("sigmet.idx.hs")

kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

head(kegg.sets.hs, 3)
```

Further data manipulation

```{r}
foldchanges = res$log2FoldChange

names(foldchanges) = res$entrez

head(foldchanges)
```



Using GAGE

```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs, same.dir = TRUE)
```

Examining the GAGE data


```{r}
attributes(keggres)
```

```{r}
head(keggres$greater)
```
 
```{r}
head(keggres$less)
```
 
Applying labels to the data 

```{r, cache=TRUE}
lapply(keggres, head)
```

Examining the top 5 of the upregulated pathways, then just getting the IDs

```{r}
pathways <- data.frame(id = row.names(keggres$greater), keggres$greater)

head(pathways)
```

Using the pathway package to use the pathview function in order to make a pathway plot 

```{r}
pathview(gene.data = foldchanges, pathway.id = "hsa04110")


```


Manipulating the above graph

```{r}
#pathview(gene.data = foldchanges, pathway.id = "hsa04110", kegg.native = FALSE)
```

Pulling out the top five upregulated pathways 

```{r}
keggrespathways <- rownames(keggres$greater) [1:5]

#finding their IDs

keggresids <- substr(keggrespathways, start = 1, stop = 8)

keggresids
```

Creating plots of the top 5 upregulated pathways 

```{r, warning=FALSE}
pathview(gene.data = foldchanges, pathway.id = keggresids, species = "hsa")
```

## Doing Gene Ontology with GAGE

```{r}
data(go.sets.hs)
data(go.subs.hs)
gobpsets <- go.sets.hs[go.subs.hs$BP]

gobpres <- gage(foldchanges, gsets = gobpsets, same.dir = TRUE)

lapply(gobpres, head)
```

## Reactome Pathway Analysis

Using reactome to analyze over-representation enrichment analysis 

```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

Making a table

```{r}
write.table(sig_genes, file = "significant_genes.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)

```


The plots

```{r}

```


