---
title: "BIMM 143 Transposon Analysis"
author: "Jacob Gerzenshtein"
date: "May 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Transposon Analysis 

Installing the relevant packages

```{r}
#commented out the install of the package so not done every time document is knit
#install.packages("tidyverse")
```

Loading relevant package libraries

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)
```

Loading the experimental file(s) for analysis 

```{r}
exp_data <- read.csv("Experimental_results.csv")

head(exp_data)
```

Getting rid of first column X 

```{r}
exp_data <- select(exp_data, -X)
head(exp_data)
```

Select function practice

```{r}
Mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)
head(Mut_str_env)
```

Getting ready for plotting using ggplot2

Have to get the table ready to be x versus y, manipulating the current data to do that 

Just flip the time and frequency into five data points 

The BC ID, Mut ID, Strain, and Environment will have to repeat for each time/frequency point

Using the gather function to do just that 

```{r}
exp_rearranged <- gather(exp_data, Generation, Frequency, H_0:H_96)
head(exp_rearranged)
```

Removing the H_ part from the generation column to make it neater

Will make two columns - the first will have H_ and the second will have the actual timepoint

```{r}
table_for_graph <- separate(exp_rearranged, Generation, into = c("H", "Time"))
head(table_for_graph)
```

Removing the H column completely because it has no effect

```{r}
table_for_graph <- select(table_for_graph, -H)
head(table_for_graph)
```

Removing the NA's from the data

```{r}
table_cleaned <- na.omit(table_for_graph)
table_cleaned$Time <- as.numeric(table_cleaned$Time)
head(table_cleaned)
```


Plotting the data

```{r}
ggplot(table_cleaned) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = 0.1, colour = "#000033") + facet_grid(.~Strain) + ggtitle("Barcode Trajectories") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Time, Hours") + ylab("Log10(Barcode Frequency)")
```

Picking one mutation to see how it behaves in different strains

```{r}
mut34 <- filter(table_cleaned, table_cleaned$Mut_ID == "34")

mut34
```

Plotting this data

```{r}
ggplot(mut34) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = 0.1, colour = "#000033") + facet_grid(.~Strain) + ggtitle("Barcode Trajectories") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Time, Hours") + ylab("Log10(Barcode Frequency)")
```

```{r}
View(table_cleaned)
```

Filtering out bad data

```{r}
mut34_f <- filter(mut34, mut34$Frequency > -5)
```

Plotting again

```{r}
ggplot(mut34_f) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = 0.1, colour = "#000033") + facet_grid(.~Strain) + ggtitle("Barcode Trajectories") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Time, Hours") + ylab("Log10(Barcode Frequency)")
```

Takingfiltered data for mutation 34 and filtering out information about one particular barcode

```{r}
BC_25361 <- filter(mut34_f, mut34_f$BC_ID == "25361", mut34_f$Strain == "DivAnc")

BC_25361
```

Plotting
```{r}
ggplot(BC_25361) + geom_point(aes(x = Time, y = Frequency, group = BC_ID), alpha = 0.9, colour = "#000033") + facet_grid(.~Strain) + ggtitle("Trajectory") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Time, Hours") + ylab("Log10(Frequency)")
```

Fitting the line to the plot 

```{r}
BC_plot <- ggplot(BC_25361) + 
  aes(x = Time, y = Frequency, group = BC_ID) + 
  geom_point(alpha = 0.9, colour = "#000033") + 
  facet_grid(.~Strain) + ggtitle("Trajectory") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  xlab("Time, Hours") + ylab("Log10(Frequency)")

BC_plot_lm <- BC_plot + geom_smooth(se = FALSE, method = "lm")

BC_plot_lm
```

Regressioon statistics

```{r}
#goes y to x instead of x to y
regression_model = lm(Frequency~Time, BC_25361)
summary_data = summary(regression_model)
summary_data

Time <- summary_data$coefficients[2]

Time


Intercept <- summary_data$coefficients[1]

Intercept
```

